### 2019-03-22 IEEE_TC_submission.d/symmetry_properties.tex
ssr

\begin{corollary}\label{cor:ncf_not_rsymm}
For every $n \geq 2$, there is an $n$-variable NCF that is not $n-1$ symmetric.
\end{corollary}

\begin{corollary}\label{cor:ncf_r_symmetric_layers}
(i) An NCF with a default-normalized nested canalyzing representation consisting
of $q$ layers is $2q$-symmetric, and is not $(q-1)$-symmetric.
(ii) An $r$-symmetric NCF has a default-normalized nested canalyzing representation
with at most $r-1$ layers, and has an unnormalized nested canalyzing
representation with at most $r$ layers.  
\end{corollary}

@@

Let $f(x_1, x_2, \ldots, x_n)$ be an $r$-symmetric Boolean function of $n$
variables.
We assume that the symmetry groups $g_i$, $1 \leq i \leq r$,
of $f$ are given.
Let $m_i$ denote the number of variables
in the group containing variable $x_i$, $1 \leq i \leq m$.
Thus, in the group containing $x_i$, the number
of variables which can take on the value 1 varies from 0
to $m_i$, $1 \leq i \leq n$.
We also assume that $f$ is
given as a table $T$ with $\mu = \prod_{i=1}^r (m_i+1)$ rows.
Each row of $T$ specifies an $r$-tuple $(c_1, c_2, \ldots, c_r)$,
where $c_i$ is the number of variables of group $g_i$ which
have the value 1, along with the \{0,1\} value of $f$ for that $r$-tuple.
As will be seen, our algorithm for determining whether $f$ is
an NCF runs in $O(r^2\mu)$ time.
The algorithm relies on the following observation.

\begin{observation}\label{obs:rsym_canalyzing}
Suppose $f(x_1, x_2, \ldots, x_n)$ is an $r$-symmetric Boolean function.
Consider any variable $x_i$ and suppose
the number of variables in the group
containing $x_i$ is $m_i$, $1 \leq i \leq r$.
\begin{enumerate}
\item The function $f$ is canalyzing in variable $x_i$
with canalyzing value 1 iff all the table entries where the number
of 1's in the group containing $x_i$ is nonzero have the same value
of $f$. (This value of $f$ is the canalyzed value when $x_i = 1$.)
\item The function $f$ is canalyzing in variable $x_i$ with
canalyzing value 0 iff all the table entries where the number of
1's in the group containing $x_i$ is less than $m_i$ have the same
value of $f$. (This value of $f$ is the canalyzed value when $x_i = 0$.)
\end{enumerate}
\end{observation}
We now explain how Observation~\ref{obs:rsym_canalyzing} can be
used to develop an iterative algorithm for determining whether 
$f$ is an NCF, and if so, constructing a default-normalized 
representation (defined in Section~\ref{sec:prelim}) for $f$.
At beginning of iteration $j$, the algorithm uses a set $X_j$ of remaining variables,
table $T_j$ with $\mu_j$ rows and the number of remaining groups is $g_j$.
Initially (i.e., $j = 1$), $X_1$ = $X$ = $\{x_1, \ldots, x_n\}$,
$T_1 = T$ (the given table $T$ for $f$), 
$\mu_1 = \mu$ (the number of rows of $T$), $g_j = r$ (the number of
symmetry groups of $f$) and the NCF representation for $f$ is empty.
The algorithm carries out iteration $j$ if $g_j \geq 1$; 
in that iteration, the algorithm performs Steps I, II and III
as described below.

@@

To estimate the running time, we note that in iteration $j$, we can
determine in $O(g_j \mu_j)$ = $O(r \mu)$ time using 
Observation~\ref{obs:rsym_canalyzing}
whether there is a variable that is canalyzing.
(This can be done by considering one arbitrary variable 
from each group.)
Once a variable is identified, the other steps in that iteration
can be completed in $O(\mu_j)$ = $O(\mu)$ time.
Thus, the time for each iteration is $O(r \mu)$.
Since there are at most $r$ iterations, the running time of the
algorithm is $O(r^2 \mu)$.
The following theorem summarizes this result.

>>>>

\begin{corollary}\label{cor:ncf_not_rsymm}
For every $n \geq 2$, there is an $n$-variable NCF that is not $n-1$ symmetric. \QED
\end{corollary}

\begin{corollary}\label{cor:ncf_r_symmetric_layers}
(i) An NCF with a default-normalized nested canalyzing representation consisting
of $q$ layers is $2q$-symmetric, and is not $(q-1)$-symmetric.
(ii) An $r$-symmetric NCF has a default-normalized nested canalyzing representation
with at most $r-1$ layers, and has an unnormalized nested canalyzing
representation with at most $r$ layers.   \QED
\end{corollary}

@@

Let $f(x_1, x_2, \ldots, x_n)$ be an $r$-symmetric Boolean function of $n$
variables.
We assume that the symmetry groups $g_i$, $1 \leq i \leq r$,
of $f$ are given.
Let $m_i$ denote the number of variables
in the group containing variable $x_i$, $1 \leq i \leq m$.
Thus, in the group containing $x_i$, the number
of variables which can take on the value 1 varies from 0
to $m_i$, $1 \leq i \leq n$.
We also assume that $f$ is
given as a table $T$ with $\mu = \prod_{i=1}^r (m_i+1)$ rows.
Each row of $T$ specifies an $r$-tuple $(c_1, c_2, \ldots, c_r)$,
where $c_i$ is the number of variables of group $g_i$ which
have the value 1, along with the \{0,1\} value of $f$ for that $r$-tuple.
As will be seen, our algorithm for determining whether $f$ is
an NCF runs in $O(r\mu)$ time.
The algorithm relies on the following observation.

\begin{observation}\label{obs:rsym_canalyzing}
Suppose $f(x_1, x_2, \ldots, x_n)$ is an $r$-symmetric Boolean function.
Consider any variable $x_i$ and suppose
the number of variables in the group
containing $x_i$ is $m_i$, $1 \leq i \leq r$.
\begin{enumerate}
\item The function $f$ is canalyzing in variable $x_i$
with canalyzing value 1 iff all the table entries where the number
of 1's in the group containing $x_i$ is nonzero have the same value
of $f$. (This value of $f$ is the canalyzed value when $x_i = 1$.)
\item The function $f$ is canalyzing in variable $x_i$ with
canalyzing value 0 iff all the table entries where the number of
1's in the group containing $x_i$ is less than $m_i$ have the same
value of $f$. (This value of $f$ is the canalyzed value when $x_i = 0$.)
\QED
\end{enumerate}
\end{observation}
We now explain how Observation~\ref{obs:rsym_canalyzing} can be
used to develop an iterative algorithm for determining whether 
$f$ is an NCF, and if so, constructing a default-normalized 
representation (defined in Section~\ref{sec:prelim}) for $f$.
At beginning of iteration $j$, the algorithm uses a set $X_j$ of remaining variables,
table $T_j$ with $\mu_j$ rows and the number of remaining groups is $g_j$.
Initially (i.e., $j = 1$), $X_1$ = $X$ = $\{x_1, \ldots, x_n\}$,
$T_1 = T$ (the given table $T$ for $f$), 
$\mu_1 = \mu$ (the number of rows of $T$), $g_j = r$ (the number of
symmetry groups of $f$) and the NCF representation for $f$ is empty.
The algorithm carries out iteration $j$ if $g_j \geq 1$; 
in that iteration, the algorithm performs Steps I, II and III
as described below.

@@

To estimate the running time, we note that in iteration $j$, we can
determine in $O(g_j \mu_j)$ = $O(r \mu_j)$ time using 
Observation~\ref{obs:rsym_canalyzing}
whether there is a canalyzing variable.
(This can be done by considering an arbitrary variable 
from each group.)
Once a variable is identified, the other steps in that iteration
can be completed in $O(\mu_j)$ time.
Thus, the time for iteration $j$ is $O(r \mu_j)$.
We note that in each iteration, the number of rows in the table
is reduced by at least a factor of 2; that is, $\mu_{j+1} \leq \mu_j/2$.
To see this, suppose the canalyzing variable $x_i$ found in iteration $j$
is in group $\alpha_p$, which has $m_p$ variables. 
Thus, in $T_j$, there are $m_p+1 \geq 2$ possibilities for the number of 1's 
in $\alpha_p$. 
As indicated in Substep~5 of Step~II above, in $T_{j+1}$, 
we only retain those rows of $T_j$ where all the variables 
in $\alpha_p$ have the same value.
In other words, there is only one possibility for the number of
1's in group $\alpha_p$.
Thus, $\mu_{j+1}$, the number of rows in $T_{j+1}$,
is at most $\mu_j/(m_p+1)$ $\leq$ $\mu_j/2$.
It follows by simple induction that $\mu_j \leq \mu/2^{j-1}$.
Thus, the running time over all the $r$ iterations is bounded by
$O(r [\sum_{j=1}^{r}(\mu/2^{j-1})])$ = $O(r\mu)$.
The following theorem summarizes this result.
